Template Master's Thesis

### 2023-08-05 paper.tex
Update introduction

In public permissionless blockchains such as Bitcoin \citep{nakamotoBitcoin2008} and Ethereum \citep{buterin2014ethereum}, individuals can participate with multiple pseudonymous addresses. Any user can create a large number of these addresses at virtually no cost. Also, contrary to popular belief, permissionless blockchains are completely transparent. All confirmed transactions are publicly observable and stored as part of the blockchain’s history. 
This has opened up a nascent scientific field dealing with entity identification within blockchain networks. By analyzing on-chain data and detecting usage patterns, researchers try to cluster addresses that are controlled by the same user. Most of these methods are heuristics, as the lack of a ground truth makes evaluation hard \newline
On the one hand, knowing which addresses belong to the same real-world entity is advantageous, as it allows for a better assessment of network properties in terms of usage, wealth distribution, and fraudulent activity \citep{FV:17}. For example, a user could manipulate a seemingly fair on-chain voting process by splitting his voting power among multiple addresses. \newline
On the other hand, the lack of privacy is also detrimental to most financial use cases(financial applications). In response to this, several privacy-enhancing protocols and tools have been proposed to obfuscate transaction tracing. The most prominent/widely used of which is Tornado cash. This reaction has, in turn, sparked the development of other clustering heuristics exploiting user behavior interacting with privacy-enhancing protocols. \newline

%Privacy will be an important topic in the future.

%Identifying addresses that belong to the same entity is predominantly done through heuristics.

\textbf{Related work}\\
Previous work in this field can be broadly distinguished between heuristics developed for the unspent transaction output (UTXO) model (e.g Bitcoin, ZCash) and the account-based model (e.g. Ethereum, Polygon). While both share the concept of addresses, the concept of accounts is not present in UTXO-based blockchains and the way in which transactions are processed is fundamentally different. Therefore, clustering heuristics are not applicable to account-model blockchains. Meiklejohn et al. proposes

\textbf{UTXO-model clustering heuristics} \\%not the focus of this work and therefore not explained
A number of address clustering heuristics in the UTXO setup have been proposed Bitcoin (Meiklejohn), Litecoin, and ZCash. Satoshi proposed using a new address for each transaction and most wallets work this way.

In contrast to the UTXO-model, in account-based blockchains, native transactions can only move funds between a single sender and receiver (payment transaction), and the “change” remains in the sender account. Subsequent transactions necessarily use the same address again. The account-based model essentially relies on address reuse on the protocol level. (Citation)

\textbf{Account-model clustering heuristics}\\
\cite{FV:17} was the first to introduce clustering heuristics for account-based blockchains. He proposed and applied heuristics that exploit patterns related to deposit addresses, multiple participation in airdrops, and token authorization mechanisms. \newline
\cite{Beres2020} argue that Victor’s heuristics, while powerful, assume participation in certain on-chain events and therefore propose more universally applicable methods. Their technique utilizes graph representation learning, a subset of machine learning interpreting transactions or token transfers as network graphs, with addresses as nodes and transactions or token transfers as edges. They further contribute to the field by proposing time-of-day activity and transaction fee-based user profiling techniques. Using ENS address pairs as ground truth, they test the methods rigorously and apply their findings to significantly reduce the privacy guarantees of Tornado Cash. They also provide a comprehensive overview and comparison of different node embedding algorithms. \newline
\cite{wu2022tutela} extend on one of \cite{Beres2020} node embedding algorithms and apply it at scale. Further, they propose five clustering heuristics in the context of Tornado Cash. They build an application built on those heuristics to report the true anonymity of an Ethereum address. All methods will be discussed in greater detail in Section 3.
Broadening the scope to entity identification, \cite{victorlüders2019} undertook an analysis of the top Ethereum-ERC20 token networks from a graph perspective. Similarly, \cite{casalebrunet2021} extend the analysis to various NFT (ERC721) ecosystems. Both find ... Payette schwager propose a segmentation of Ethereum address space into four distinct behaviour groups based on a broad range of address features (k-means).

Instead of on-chain data, \cite{yu2023} propose a new correlation method by utilizing nodes as entities, exploiting network information. This approach comes with a whole new set of limitations and problems, but one can possibly avoid the impact of privacy-enhancing technologies. This approach may be interesting once privacy-enhancing techniques are more widely used.

%network level/off-chain data ->
%transaction level/ on-chain data ->
%In the UTXO model, each transaction has to reference an output from a previous transaction. When spending a Bitcoin, a user creates a UTXO. Each UTXO incorporates a spending condition, which can only be fulfilled by the . Transactions can thought of as large graphs of inputs and outputs and most wallet implementation use a new Bitcoin address/public key for each transaction.

@@

We test the applicability of existing heuristics for our case and evaluate their success in detecting (given there is a ground truth). We estimate how many real-world entities are represented by these addresses. To achieve this, we adapt and extend existing methods to our specific situation/setting.

Propose own heuristics approaches for detecting address clusters.

@@

\textbf{Structure of the paper}\\
The remainder of this paper is structured as follows: In Section 2, we provide a brief overview of the key concepts (basics, necessary for understanding the setup), including Ethereum accounts and the EVM, tokens, Decentraland and privacy-enhancing protocols. In Section 3, we describe our data collection methodology and the data preparation. Section 4  describes and different clustering heuristics in detail. Section 5 applies the clustering heuristics from 4. Finally we discuss the results of our analysis and summarize our findings in Sections 5 and 6.

@@

\textbf{Ethereum, Polygon, EVM}\\

\textbf{Transactions, Transfer Events}\\

@@

\textbf{Metaverse, Decentraland} \\

\textbf{Privacy-enhancing protocols, mixing, Tornado} \\

@@

The initial dataset contained 473,927 unique Web3 addresses. As we are only interested in clustering methods leveraging on-chain transaction or token transfer data, we omit addresses that were not recorded on either Ethereum or Polygon. This was achieved using data/information from Blockscan\footnote{\url{https://blockscan.com/}}. In total, 59,651 addresses were recorded on Ethereum, 129,988 addresses on Polygon and 52,095 on both networks simultanously. Therefore, the remaining address set comprises 137,544 addresses. Figure 1 visualizes the subsets in a VENN diagram. 
We used these addresses to gather/collect transaction and token transfer receipts from Ethereum Mainnet and Polygon PoS. Ethereum or Polygon data can be accessed directly through a node or an application programming interface (API) provider such as Infura or Etherscan. The absence of account indexing in Ethereum or Polygon makes queries of the type "Retrieve all past transactions and token transfers of address \texttt{0xABC}" challenging, as one would have to scan through every block and every token transfer event emitted by a specified set of token contracts. Fortunately, Etherscan offers/provides an API Endpoint Module for ``Accounts`` that facilitates the retrieval of transactions and token transfer events for a given address. Using our network specific address subsets, we were able to significantly reduce the amount of API calls.

We gathered all transactions and token transfer data of the addresses until block 17,670,000 on Ethereum (July 11, 2023) and block 44,990,000 on Polygon (July 12, 2023). The results were stored in comma-separated value (CSV) files and loaded into a MongoDB database\footnote{\url{https://www.mongodb.com/}}.

>>>>

Public permissionless blockchains such as Bitcoin \citep{nakamotoBitcoin2008} and Ethereum \citep{buterin2014ethereum} allow users/individuals to participate with multiple pseudonymous addresses. The creation of these addresses is virtually cost free. Contrary to popular belief, these blockchains are entirely transparent. All transactions are stored as part of the blockchain's history and publicly observable.
This has opened up a nascent scientific field dealing with entity identification within blockchain networks. Researchers try to cluster addresses controlled by the same user by analyzing on-chain data and detecting usage patterns. The majority of these methods are heuristic which often makes them difficult to evaluate due to the absence of any ground truth. \newline
On one hand, identifying the addresses that belong to the same real-world entity is beneficial as it allows better evaluation of network properties with respect to usage, distribution of wealth and detecting fraudulent activities \citep{FV:17}. For instance, if a user distributes their voting power across various addresses, they might manipulate an on-chain voting process that seems fair at the outset. \newline
Conversely, the lack of privacy is also detrimental to most financial use cases (DeFi). As a response, multiple privacy-enhancing tools and protocols have been proposed to obfuscate transaction tracing. 
Nonetheless, these protocols are not yet widely adopted and careless user behaviour can undermine the privacy guarantees they offer. \newline
The topic of privacy, or lack thereof, will remain significant for future blockchain development and research.

%However, they are not widely used even cannot guarantee perfect privacy (factors: size of anonymity set, usage /user behaviour of protocol, privacy guarantees depend on other users).
%The most prominent/widely used non-custodial mixer is Tornado cash (Schär, Tutela).
%This reaction has, in turn, sparked the development of other clustering heuristics exploiting user behavior interacting with privacy-enhancing protocols.

\textbf{Related work}\\
Previous work in this field can be broadly distinguished between heuristics developed for the unspent transaction output (UTXO) model (e.g Bitcoin, ZCash) and the account-based model (e.g. Ethereum, Polygon). While both models share the concept of addresses, the notion of accounts is not present in UTXO-based blockchains. The way in which transactions are processed is fundamentally different, and therefore, clustering heuristics are not applicable to both paradigms.
%\textbf{UTXO-model clustering heuristics} 
A number of address clustering heuristics in the UTXO setup have been proposed for Bitcoin and derivatives \citep{Androulaki2013, Meiklejohn2013, Haslhofer2016, jourdan2018, kappos2022}. All these methods will not be the focus of this work and therefore not discussed. As suggested by Nakamoto in the Bitcoin whitepaper, most Bitcoin wallet implementations use a new key pair for each transaction to keep them from being linked to a common owner. \newline
In contrast to the UTXO-model, native transactions in account-based blockchains can only move funds between a single sender and receiver, and the ``change``remains in the sender account. Subsequent transactions necessarily use the same address again. The account-based model essentially relies on address reuse on the protocol level. \citep{Beres2020} \newline
% Make an assumption that people only use a small set of addressesyy
%\textbf{Account-model clustering heuristics}
Clustering heuristics for account-based blockchains were first introduced by \cite{FV:17}. He proposed and applied heuristics that exploit patterns related to deposit addresses, multiple participation in airdrops, and token authorization mechanisms. \newline
\cite{Beres2020} propose more universally applicable methods, as they argue that Victor’s heuristics, while powerful, assume participation in certain on-chain events. Their technique interprets transactions or token transfers as network graphs, with addresses as nodes and asset transfers as edges. They quantitavely compare graph-representation learning algorithms (a subset of machine learning) and propose further user profiling techniques based on time-of-day activity and transaction fees. Using ENS address pairs as ground truth (quasi-identifiers), the authors rigorously test their methods and apply their findings to significantly reduce the privacy guarantees of Tornado Cash. \newline
\cite{wu2022tutela} extend on one of \cite{Beres2020} graph representation learning (node-embedding) algorithms and apply it at scale. Further, they also propose a set of new clustering heuristics targeting Tornado Cash, showing that careless user behavior can still reveal identity. They build an application built on those heuristics to measure the anonymity of an Ethereum address. \newline
All of the methods mentioned will be discussed in greater detail in Section 3. \newline
Broadening the scope of entity identification, \cite{victorlüders2019} analyzed the top Ethereum ERC-20 token networks from a graph perspective. Similarly, \cite{casalebrunet2021} apply a similar analysis to various non-fungible token (NFT) ecosystems. Both find similarities to interaction graphs in social networks (hub-and-spoke). \cite{Payette2017} propose a segmentation of Ethereum address space into four distinct behaviour groups sharing similar attributes using k-means clustering.

\cite{yu2023} propose a novel approach for correlation analysis by exploiting network information, rather than treating users as entities and analyzing on-chain data. Although this approach has the potential to avoid the impact of privacy-enhancing technologies, it introduces a new set of limitations and problems. Thus, this approach may be of great interest for further research, particularly when privacy-enhancing techniques become more widely used.

%network level/off-chain data ->
%transaction level/ on-chain data ->
%In the UTXO model, each transaction has to reference an output from a previous transaction. When spending a Bitcoin, a user creates a UTXO. Each UTXO incorporates a spending condition, which can only be fulfilled by the . Transactions can thought of as large graphs of inputs and outputs and most wallet implementation use a new Bitcoin address/public key for each transaction.
%In addition, the authors provide a comprehensive overview and comparison of various node embedding algorithms.

@@

We test the applicability of existing heuristics for our case and evaluate their success in detecting (given there is a ground truth). We estimate how many real-world entities are represented by these addresses. To achieve this, we adapt and extend existing methods to our specific situation/setting. Propose own heuristics approaches for detecting address clusters.

@@

The remainder of this paper is structured as follows: In Section 2, we provide a brief overview of the key concepts (basics, necessary for understanding the setup), including Ethereum accounts and the EVM, tokens, Decentraland and privacy-enhancing protocols. In Section 3, we describe our data collection methodology and the data preparation. Section 4  describes and different clustering heuristics in detail. Section 5 applies the clustering heuristics from 4. Finally we discuss the results of our analysis and summarize our findings in Sections 5 and 6.

@@

Ethereum Mainnet and Polygon PoS are both account-model blockchains and use the same execution logic, the Ethereum Virtual machine.

- Polygon PoS is a so-called sidechain, closely associated with Ethereum mainnet and used for scaling

%\textbf{Ethereum, Polygon, EVM}\\

\textbf{Transactions and Transfer Events}\\

@@

\textbf{Metaverse and Decentraland} \\
- Much of the Decentraland infrastructure has been deployed on Ethereum Mainnet

\textbf{Privacy-enhancing protocols} \\

@@

The original dataset comprised 473,927 distinct (Web3) addresses. As we only focus on clustering methods using on-chain transaction or token transfer data, we exclude addresses that have not been recorded on either Ethereum or Polygon. To accomplish this, we used data from Blockscan\footnote{\url{https://blockscan.com/}}. Overall, 59,651 addresses were recorded on Ethereum, 129,988 on Polygon, with 52,095 addresses appearing on both networks simultaneously. The set of remaining addresses consists of 137,544 addresses. Subsets are visualized in a Venn diagram in Figure 1. 
These addresses were used to collect transactions and token transfer events from Ethereum Mainnet and Polygon PoS. Ethereum or Polygon data can be accessed directly through a node or an application programming interface (API) provider like Infura or Etherscan. The absence of account indexing in Ethereum or Polygon poses a challenge for retrieving all past transactions and token transfers of a specific address, as it requires scanning through all blocks and token transfer events emitted by designated token contracts. Fortunately, Etherscan offers/provides an API Endpoint Module for ``Accounts`` that facilitates the retrieval of transactions and token transfer events for a given address. By using our network-specific address subsets, we were able to significantly decrease the number of required API calls. 
We gathered all transaction and token transfer data of the addresses up until block 17,670,000 on Ethereum (July 11, 2023) and block 44,990,000 on Polygon (July 12, 2023). The output was saved in comma-separated value (CSV) files and imported into a MongoDB database\footnote{\url{https://www.mongodb.com/}}.
